Tree establishment in models of forest dynamics: a key priority for further research

### 2023-03-12 manuscript/manuscriptRW.Rmd
change deffinition for forclim 11

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree establishment [@käber2022]. This is a challenging experiment for
these models as they do also have to estimate the forest stand above the
tree establishment. By tree establishment, we refer to the passing of
trees across a specific diameter threshold ("ingrowth"). We included
models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @mette2014; @lexer2001]. Because of the large variability
in tree establishment patterns in nature and the large number of factors
driving this process --- including some that are not incorporated in
most models such as deer browsing --- we do not aim at a detailed
statistical evaluation of each model. Instead, we aim at evaluating the
general ingrowth patterns and magnitudes simulated by the models, and
benchmark the simulated regeneration niche of multiple species against
empirical data along a wide environmental gradient of temperature,
moisture, and light availability.

@@

Two models, ForClim and ForCEEPS, were used here with two alternative
variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
module that adheres closely to the concept introduced by @botkin1972,
whereas ForClim variant 11 is adopting a rather different approach that
considers space filling along with different 'points of entry' for the
action of specific environmental factors [@huber2020]. These two model
variants allow us to evaluate a more process-based module (variant 11)
against a simple module (variant 1), while the rest of the model
structure is identical. Similarly, the two ForCEEPS variants allow us to
isolate the importance of the canopy feedback (i.e. simulated actual
composition and relative abundance of species in the plot) via seed
trees for the quantity and quality of simulated ingrowth, as one variant
includes this feedback whereas the other does not.

>>>>

In the present study, for the first time, a large number of forest
models commonly used to assess forest dynamics under climate change are
evaluated against a continental-scale, multi-species harmonized dataset
on tree establishment [@käber2022]. This is a challenging experiment for
these models as they do also have to estimate the forest stand above the
tree establishment. By tree establishment, we refer to the passing of
trees across a specific diameter threshold ("ingrowth"). We included
models that are based on a range of "philosophies", from models
operating at the stand to the global scale as well as the range of
models from empirically derived to "process-based" [e.g., @bugmann1996;
@fabrika2012; @lexer2001; @reyer2014]. Because of the large variability
in tree establishment patterns in nature and the large number of factors
driving this process --- including some that are not incorporated in
most models such as deer browsing --- we do not aim at a detailed
statistical evaluation of each model. Instead, we aim at evaluating the
general ingrowth patterns and magnitudes simulated by the models, and
benchmark the simulated regeneration niche of multiple species against
empirical data along a wide environmental gradient of temperature,
moisture, and light availability.

@@

Two models, ForClim and ForCEEPS, were used here with two alternative
variants. ForClim variant 1 [@bugmann1996] is based on a recruitment
module that adheres closely to the concept introduced by @botkin1972,
whereas ForClim variant 11 is adopting a different approach where
individual species properties and its relation to the environment are
more important [@huber2020]. These two model variants allow us to
evaluate a more process-based and complex module (variant 11) against a
simple module (variant 1), while the rest of the model structure is
identical. Similarly, the two ForCEEPS variants allow us to isolate the
importance of the canopy feedback (i.e. simulated actual composition and
relative abundance of species in the plot) via seed trees for the
quantity and quality of simulated ingrowth, as one variant includes this
feedback whereas the other does not.
